The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-09 rational.tex
added wire over the multiplicative algebra

Now consider the following algebraic manipulation (assuming $a \neq 0$):
\[\begin{array}{rclll}
&& a & \mbox{initial~$a$} & (0) \\
&=& a * 1 & \mbox{$1$~is~identity~for~$*$} & (1) \\
&=& a * ((1/a) * a) & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (2) \\
&=& (a * (1/a)) * a & \mbox{$*$~is~associative} & (3) \\
&=& 1 * a & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (4) \\
&=& a & \mbox{$1$~is~identity~for~$*$} & (5)
\end{array}\]
As the semantics in Sec.~\ref{sec:rat} formalizes, the witness to this 
{{a <-> a}} isomorphism computes as follows:
\begin{itemize}
\item We start at line (0) with the value {{20}};
\item We proceed to line (1) with the value {{(20,())}};
\item We proceed to line (2) with the value {{(20,((1/alpha),alpha))}} where
  $\alpha$ is essentially a logic variable;
\item We proceed to line (3) with the value {{((20,1/alpha),alpha)}};
\item At line (4) we require that $20$ and $1/\alpha$ cancel each other which
  forces the unification $\alpha=20$, i.e., the value becomes
  {{((20,1/20),20)}} which simplifies to {{((),20)}};
\item We reach line (5) with the value {{20}}.
\end{itemize}

>>>>

Now consider the following algebraic manipulation (assuming $a \neq 0$):
%% \[\begin{array}{rclll}
%% && a & \mbox{initial~$a$} & (0) \\
%% &=& a * 1 & \mbox{$1$~is~identity~for~$*$} & (1) \\
%% &=& a * ((1/a) * a) & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (2) \\
%% &=& (a * (1/a)) * a & \mbox{$*$~is~associative} & (3) \\
%% &=& 1 * a & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (4) \\
%% &=& a & \mbox{$1$~is~identity~for~$*$} & (5)
%% \end{array}\]

%% \[\begin{array}{rcllllll}
%%  \!&\! a \!&\! \!&\!        \!&\! \!&\!    \!&\! \mbox{initial~$a$} \!&\! (0) \\
%% =\!&\! a \!&\!*\!&\!        \!&\!1\!&\!    \!&\! \mbox{$1$~is~identity~for~$*$} \!&\! (1) \\
%% =\!&\! a \!&\!*\!&\! ((1/a) \!&\!*\!&\! a) \!&\! \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} \!&\! (2) \\
%% =\!&\!(a \!&\!*\!&\! (1/a)) \!&\!*\!&\! a  \!&\! \mbox{$*$~is~associative} \!&\! (3) \\
%% =\!&\!   \!&\!1\!&\!        \!&\!*\!&\! a  \!&\! \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} \!&\! (4) \\
%% =\!&\!   \!&\! \!&\!        \!&\! \!&\! a  \!&\! \mbox{$1$~is~identity~for~$*$} \!&\! (5)
%% \end{array}\]

\begin{center}
\scalebox{1.1}{
\includegraphics{diagrams/thesis/algebra-wire2.pdf}
}
\end{center}

As the semantics in Sec.~\ref{sec:rat} formalizes, the witness to this 
{{a <-> a}} isomorphism computes as follows:
\begin{itemize}
\item We start at line (0) with the value {{20}};
\item We proceed to line (1) with the value {{(20,())}};
\item We proceed to line (2) with the value {{(20,((1/alpha),alpha))}} where
  $\alpha$ is essentially a logic variable;
\item We proceed to line (3) with the value {{((20,1/alpha),alpha)}};
\item At line (4) we require that $20$ and $1/\alpha$ cancel each other which
  forces the unification $\alpha=20$, i.e., the value becomes
  {{((20,1/20),20)}} which simplifies to {{((),20)}};
\item We reach line (5) with the value {{20}}.
\end{itemize}
